---
title: "Importing and Cleaning data"
author: "Huruy Belay"
date: "May 19, 2017"
output: html_document
---

## Using readxl

##### Required libraries

```{r, results='hide'}
library(readxl)
library(tidyr)
```

##### Importing the data

```{r, results='hide'}
mbta <- read_excel("mbta.xlsx", skip = 1)
head(mbta)
```

* **None**: skiped the first raw of the data

##### Examining the data

```{r, results='hide'}
str(mbta)
head(mbta)
summary(mbta)
```

##### Removing unnecessary rows and columns

```{r}
mbta1 <- mbta[-c(1,7,11), -1]
head(mbta1)
```

**Note:**
     * Removed rows 1,7 and 11.
     * Removed column 1.

##### Observations are stored in columns

If you see the data there are observations are stored in columns and we should fix this using the gather function from the tidyr package.

```{r}
mbta2 <- gather(mbta1, month, thou_riders, -mode)
head(mbta2)
```

At this time thins are ok nice. However, we should check the structure of the variables. The thou_riders variable should be numeric.

```{r}
mbta2$thou_riders <- as.numeric(mbta2$thou_riders)
```

##### Variables are stored in both rows and columns

If you look the data again, variables are stored in both rows and columns. We should fix this also using one of the tidyr packages.

```{r}
mbta3 <- spread(mbta2, mode, thou_riders)
head(mbta3)
```

The data looks better, but one thing should fix. We should separate the variable 'r month' in to year and month.

```{r}
mbta4 <- separate(mbta3, month, c("year", "month"))
head(mbta4)
```

Finally we can start analysis. But alwayse before analysis, it is good to screen our data using some of the data sceenings.

```{r}
summary(mbta4)

hist(mbta4$Bus)
```


**Note:** I used this data from the Datacamp.